Structural Breaks in Inflation Dynamics within the European Monetary Union

### 2011-05-12 pkg/paper/Paper/draft.Rnw
new draft with fixmes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@86 b679b2f7-6539-45d0-a605-5ddcb976cde9

## additional functions
momentgen <- function(flag=1,x, start=yearmon(c(1990,1998,2007)), end=yearmon(c(1998+11/12,2007+11/12,2010+11/12))) {
# 1:mean, 2:variance, 3:skewness
if(flag==1){
mean = mat.or.vec(ncol(x),length(start))
for(jj in 1:length(start)){
x1 <- window(x, start=start[jj], end=end[jj])
for(i in 1:ncol(x1)) {
mean[i,jj] = mean(na.omit(x1[,i]))
}
}
ret1 <- c(quantile(mean[,1], probs=.25), quantile(mean[,1], probs=.50), quantile(mean[,1], probs=.75))
ret1 <- cbind(ret1)
ret2 <- c(quantile(mean[,2], probs=.25), quantile(mean[,2], probs=.50), quantile(mean[,2], probs=.75))
ret2 <- cbind(ret2)
ret3 <- c(quantile(mean[,3], probs=.25), quantile(mean[,3], probs=.50), quantile(mean[,3], probs=.75))
ret3 <- cbind(ret3)
ret <- cbind(ret1,ret2,ret3)
colnames(ret) <- c(start[1], start[2], start[3])
rownames(ret) <- c("lower", "medium", "upper")
return(ret)
}
if(flag==2){
var = mat.or.vec(ncol(x),length(start))
for(jj in 1:length(start)){
x1 <- window(x, start=start[jj], end=end[jj])
for(i in 1:ncol(x1)) {
var[i,jj] = var(na.omit(x1[,i]))
}
}
ret1 <- c(quantile(var[,1], probs=.25), quantile(var[,1], probs=.50), quantile(var[,1], probs=.75))
ret1 <- cbind(ret1)
ret2 <- c(quantile(var[,2], probs=.25), quantile(var[,2], probs=.50), quantile(var[,2], probs=.75))
ret2 <- cbind(ret2)
ret3 <- c(quantile(var[,3], probs=.25), quantile(var[,3], probs=.50), quantile(var[,3], probs=.75))
ret3 <- cbind(ret3)
ret <- cbind(ret1,ret2,ret3)
colnames(ret) <- c(start[1], start[2], start[3])
rownames(ret) <- c("lower", "medium", "upper")
return(ret)
}
if(flag==3){
skewness= mat.or.vec(ncol(x),length(start))
for(jj in 1:length(start)){
x1 <- window(x, start=start[jj], end=end[jj])
for(i in 1:ncol(x1)) {
skewness[i,jj] = skewness(na.omit(x1[,i]))
}
}
ret1 <- c(quantile(skewness[,1], probs=.25), quantile(skewness[,1], probs=.50), quantile(skewness[,1], probs=.75))
ret1 <- cbind(ret1)
ret2 <- c(quantile(skewness[,2], probs=.25), quantile(skewness[,2], probs=.50), quantile(skewness[,2], probs=.75))
ret2 <- cbind(ret2)
ret3 <- c(quantile(skewness[,3], probs=.25), quantile(skewness[,3], probs=.50), quantile(skewness[,3], probs=.75))
ret3 <- cbind(ret3)
ret <- cbind(ret1,ret2,ret3)
colnames(ret) <- c(start[1], start[2], start[3])
rownames(ret) <- c("lower", "medium", "upper")
return(ret)
}
}

@@

format_momenteuro <- function(momeuro, digits = 6, name="Euro") {
  rval <- format(round(momeuro, digits = digits), nsmall = digits)
  rval <- cbind(name,"", matrix(paste("$", rval, "$", sep = ""), ncol = 3))
  rval[1,1] <- name
  rval[2,1] <- ""
  rval[3,1] <- ""
  rval[1,2] <- "lower"
  rval[2,2] <- "medium"
  rval[3,2] <- "upper"
  rval <- apply(rval, 1, paste, collapse = " & ")
  rval <- paste(rval, "\\\\")
  rval[length(rval)] <- paste(rval[length(rval)], "\\hline")
  return(rval)
}
get_data <- function(i) {
  rval <- do.call("merge",
    lapply(i, function(j) results[[j]]$data))
  colnames(rval) <- names(results)[i]
  rval
}

@@

This framework allows us to assess the evolution of mean, variance and skewness over time. The skewness parameter turns out to be quite important to measure changes in the long run inflation dynamics, as it changes considerably over the time of the analysis. This trend would be all but invisible using a normal distribution.  

@@

The remainder of this paper is structured as follows: Section \ref{s2} presents the data, Section \ref{s3} presents the model and the estimation techniques used, Section \ref{s4} illustrates our approach using Slovenia as an example, Section \ref{s5} presents the results and Section \ref{s6} concludes.

\clearpage
\section{Data}\label{s2}
%(... first method, than data...) \\
We employ seasonally unadjusted HICP series of 21 countries ranging from January 1990 to December 2010 that are obtained from the \citet{oecd}. Countries included are Austria,	Belgium, Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,	Slovakia, Slovenia,  Spain,	Sweden, and	United Kingdom. Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity. Cypria and Malta are not included since they are very small economies. \\
The countries in this sample can be divided into three different groups. First the euro countries (Austria, Belgium, Estonia, France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, Netherlands, Portugal, Slovakia, Slovenia and Spain).  Second, the EU members not participating in the ERM~II: Czech Republic, Hungary, Poland, United Kingdom, and Sweden. Denmark stands on its own as a member of the EU and the ERM~II, but not yet a member of the EMU. \\
%(... schlussverkäufe schon hier erwähnen, einfach cool, ja das gibts, wir haben die aber weggeschätzt; generell viel angelsächsischer/präpotenter schreiben; nur 1 weg der richtige und zwar unserer; nicht erwähnen, dass andere dinge auch noch möglich; so nach motto: mich interessierts, ich machs, andere erst am ende erwähnen...)
% greece: -for winter discounts: First 10 days of January-end of February; -for summer discounts: First 10 days of July -end od August; but has always been included
As shown in a paper by \cite{pltm}, the harmonization of the treatment of sales prices led to considerable changes in some countries (Belgium, Italy, Luxembourg and Spain); failing to take this into account may lead to considerable distortions in the estimation. In these cases, we dealt with the inclusion of the sales periods effects\footnote{This was demanded by a Commission regulation, details may be found in: \citet{eurlex} and in the Appendix \ref{dataprobs}.} into the HICP series (January 2000 for Belgium, January 2001 for Italy, January 1999 for Luxembourg and January 2001 for Spain) by estimating dummies for these month\footnote{There are two month with a significant decrease in prices (January and July), followed by significant price increases in the following months (February and August).} and subtracting them from the original time series.

@@

To monitor changes in the dynamics of the inflation rates, we follow the usual strategy employed when testing for structural breaks. First, the model is fitted for the whole sample using the generalized logistic distribution. This then enables testing for the stability of the parameter estimates. If there is evidence of significant changes, the number and the timing of the break points are estimated. \\
This framework allows to trace the evolution of mean, variance and skewness of the inflation series and thus interpreting deviations in any of these parameters as possibly due to a change in monetary policy, is made feasible. Any huge intervention or change is captured by fluctuations in the parameter stability of the series and can be quantified and related to a distinct time frame. \\
%(...discuss correlation...) \\
The correlation dynamics of the series is of less importance to this framework, which concentrates on changes in mean, variance and skewness and does not focus on autocorrelation, the use of alternative tests, as described in \cite{andreou}, is also hindered by the low number of observations, due to the monthly frequencies.\footnote{In some years the Billion Prices Project of the \cite{bpp}, which provides a much higher data frequency, could solve this problem.} 
%To provide consistent estimates of the parameter vector $\hat{\phi}$, the covariance matrix $\hat{V}^{-1/2}$ is estimated using a heteroskedasticity and autocorrelation consistent covariance estimator.

@@

with location $\theta$, scale $\sigma$ and shape $\delta$. For $\delta=1$ the distribution simplifies to the logistic distribution, for $\delta<1$ it is skewed to the left and for $\delta$$>$1 it is skewed to the right (see Appendix \ref{glscore} for more details). \\
In econometrics, the logistic distribution is often used in income distributions and growth models. This is due to  its fatter tails, which fit these kind of data somewhat better. \citet{won} develop a maximum likelihood estimator if the underlying distribution is a GL distribution. A similar GL distribution is also used in \citet{tolikas} who analyze extreme risk and value--at--risk in the German stock market, although they do not use a shape parameter. Regarding inflation rates, the GL distribution is --- to our best knowledge --- only used in context with expected inflation. \\
\citet{batchelor} use a logistic distribution (not its generalization) to model the distribution of mean expected inflation rates. \\

@@

\begin{eqnarray}
\mathit{efp}(t) ~=~ \hat{V}^{-1/2} n^{-1/2} \sum_{i=1}^{\lfloor nt \rfloor} s(y_i|\hat{\theta}, \hat{\sigma}, \hat{\delta}) \quad (0 \leq t \leq 1), \nonumber
\label{eqn2}
\end{eqnarray}
%(... jesus: thinks this is incomplete ...) \\
where $\hat{V}$ is some consistent estimator of the variance of the scores, which is used to decorrelate the series as to avoid problems due to serial correlation. Then a functional central limit theorem (FCLT) for $\mathit{efp}(\cdot)$ is employed, which converges to a 3-dimensional Brownian bridge:

@@

<<echo=FALSE>>=
        x <- results$Slovenia
        slovpval <- sctest(x$efp, functional = supLM(0.1))$p.value
@
To illustrate our method, we consider the case of Slovenia. After a period of relative instability following the independence from Yugoslavia, Slovenia was successfull in realigning its economy and introduced a number of reforms leading to a stable growth in the past years. The good economic performance made it possible for Slovenia to enter the ERM~II in June 2004 and later on to introduce the euro in January 2007. The data collected contains the period from January 1996 onwards. In the first few years, Slovenia experienced very high inflation rates, but was successfull in reducing inflation to a much lower level. \\
A glance at the time series depicted in Figure \ref{fig1} of the monthly inflation rates reveals a decline in the mean around the middle of the observed time period. The $efp(\cdot)$ starts to cross the border as early as 2001 and reaches the peak somewhere between 2003 and 2004. The p--value of the supLM test is \Sexpr{round(slovpval, 5)} and clearly indicates at least one break.  The number of breaks is calculated by means of the LWZ criteria, which chooses one break in August 2003.\footnote{For the changes in the parameters, see Table \ref{resulttable}.} In this year, Slovenia introduced a number of financial reforms and severely decreased the year on year growth in the M3 monetary aggregate. Combined, these actions were enough to ensure the successfull participation in the ERM~II a year later and the introduction of the euro two years later.

@@

%(...3 punkte darstellung noch einbauen...) \\
%(...andere struktur: 1.short introduction, 2. countries in detail (6.1), 3. summary (6.2) ...) \\
%(... jesus: opening confusing and obscure...)\\
Table \ref{resulttable} presents the break dates and the parameter values of the segments as well as the dates of entry into the ERM/ERM~II and the date of the euro introduction. We formed seven groups of countries according to the observed pattern of break points. These seven groups consist of Austria and Belgium, the Eastern European countries, the two big southern European countries, Italy and Spain, Ireland, the countries with no change in the inflation dynamics, the countries with breaks in the early 90s and the member countries of the former DM--zone (excluding Austria and Belgium). \\

@@

Another combination of countries with rather identical results are the two biggest southern European countries, Italy and Spain. Both of them experienced high rates of inflation in the early 90s. Afterwards, in the phase leading up to the monetary union, mean and variance were greatly reduced. Soon after the fixing of the exchange rates, this trend changed significantly, leading to a higher mean and a much higher variance. A possible reason for this is that the fiscal responsibility that was required for an entry into the eurozone was softened after this goal was achieved. \\

@@

The disparity of the national inflation rates is still an issue. Many researchers, like \citet{hofmann}  find a considerable amount of inflation differentials. \citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity across EMU countries in terms of average inflation and its degree of persistence. In a paper examining structural convergence of inflation rates in EU countries, \citet{palomba} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar. They find that convergence over time of inflation dynamics was only partly observable. 
In a paper studying core inflation and using an aggregated euro area inflation rate, \citet{morana} finds three regimes (roughly 1980---1984, 1984---1993 and 1993---2000) governing the core inflation rate.  \\
Concerning inflation differentials, Figure \ref{greyplot} depicts the inflation rates and a rolling mean estimation, where the mean is calculated using the first moment of a fitted GL distribution.\footnote{If more countries have the same inflation rate in a given month, the lines are overlapped and thus get darker. We wold see one black line if all countries would show the same month--on--month increase at any one time. The red lines depict the means.} We do see that many of these countries follow a rather similar path. One event clearly visible is the financial crises and the short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months. Over time, there are hints, that the inflation rates converged towards some common -- albeit broad -- mean level. For the inflation series of the whole EMU, as published by the \citet{oecd}, we see an increase in mean from \Sexpr{round(euhim1, 5)} (1990(1)--1998(12)) to \Sexpr{round(euhim2, 5)} (1999(1)--2008(8)) with a strong decrease to \Sexpr{round(euhim3, 5)} (2008(9)--2010(12)) following the financial crises. Whereas variance has increased over time from    \Sexpr{round(euhiv1, 5)} to \Sexpr{round(euhim3, 5)}. Regarding skewness, we see a clear change towards a negative skewness from \Sexpr{round(euhis1, 5)} to \Sexpr{round(euhis3, 5)}.

@@

Regarding the general trend, Table \ref{grouptable} presents lower, medium and upper (25, 50, 75) percentiles of the first three moments of the inflation series during the convergence phase (January 1990 until December 1998), the euro phase (January 1999 until August 2008) and the time after the financial crisis (since September 2008\footnote{After the insolvency of the Lehman Brothers in September 2008.} until now), for the countries in the ``core'' of the eurozone (Austria, Belgium, Finland, France, Germany, Luxembourg and Netherlands), the PIIGS countries (Portugal, Italy, Ireland, Greece and Spain) and all the others. \\
%(... jesus would further differentiate between core and southeren core (or PIIGS)...)\\

\begin{longtable}{|l|l|l|l|l|}
\hline
\multicolumn{5}{|c|}{Mean} \\
\hline
Zone & Percentile & 1990(1)--1998(12) & 1999(1)--2008(8) & 2008(9)--2010(12)\\
\hline
<<echo=FALSE, results=tex>>=
euro <- cbind(table2[[1]], table2[[2]],table2[[6]],table2[[7]],table2[[8]],table2[[13]],table2[[14]])
noneuro <- cbind(table2[[3]],table2[[4]],table2[[5]],table2[[10]],table2[[15]],table2[[17]],table2[[18]],table2[[20]],table2[[21]])
piigs <- cbind(table2[[9]], table2[[11]],table2[[12]], table2[[16]],table2[[19]])
writeLines(unlist(format_momenteuro(momentgen(1,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Coreeuro")))
@
\hline
<<echo=FALSE, results=tex>>=
writeLines(unlist(format_momenteuro(momentgen(1,piigs,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="PIIGS")))
@
\hline
<<echo=FALSE, results=tex>>=
writeLines(unlist(format_momenteuro(momentgen(1,noneuro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Others")))
@
\multicolumn{5}{|c|}{Variance} \\
\hline
Zone & Quantil & 1990(1)--1998(12) & 1999(1)--2007(12) & 2008(1)--2010(12)\\
\hline
<<echo=FALSE, results=tex>>=
writeLines(unlist(format_momenteuro(momentgen(2,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Coreeuro")))
@
\hline
<<echo=FALSE, results=tex>>=
writeLines(unlist(format_momenteuro(momentgen(2,piigs,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="PIIGS")))
@
\hline
<<echo=FALSE, results=tex>>=
writeLines(unlist(format_momenteuro(momentgen(2,noneuro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Others")))
@
\hline
\multicolumn{5}{|c|}{Skewness} \\
\hline
Zone & Quantil & 1990(1)--1998(12) & 1999(1)--2007(12) & 2008(1)--2010(12)\\
\hline
<<echo=FALSE, results=tex>>=
writeLines(unlist(format_momenteuro(momentgen(3,euro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Coreeuro")))
@
\hline
<<echo=FALSE, results=tex>>=
writeLines(unlist(format_momenteuro(momentgen(3,piigs,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="PIIGS")))
@
\hline
<<echo=FALSE, results=tex>>=
writeLines(unlist(format_momenteuro(momentgen(3,noneuro,start=yearmon(c(1990,1998,2008+8/10)), end=yearmon(c(1998+11/12,2008+7/12,2010+11/12))), digits=6, name="Others")))
@
\hline
\caption{Development of first three moments}
\label{grouptable}
\end{longtable}

Regarding the mean, we see a clear evolution towards a lower mean in all groups, where the core countries show much lower values at the median. Almost the opposite is true concerning the variance of the inflation rates of the core countries, that has increased to a much higher value than at the start of the sample, with the upper percentile during the 90s now even smaller than the lower percentile for the time after the financial crisis. The same is true for the PIIGS countries, although the core country median variance is still lower; the variance for the other countries has declined significantly. Concerning skewness, we see a clear trend in the core countries for skewness to become negative, indicating that the bulk of the values now lies right to the mean. The same is true -- although to a lesser extent -- for the PIIGS and the other countries. \\

>>>>

## additional functions
cross_country_moments <- function(countries, date,
  probs = c(0.25, 0.5, 0.75))
{
  rval <- sapply(countries, function(country) {
    nb <- results[[country]]$nbreaks
    m <- t(sapply(results[[country]]$model, moments))
    bd <- if(nb < 1) {
      end(results[[country]]$data)
    } else {
      breakdates(results[[country]]$breakpoints, breaks = nb)
    }
    m[sum(as.yearmon(date) > bd) + 1, ]
  })
  apply(rval, 1, quantile, probs = probs)
}

@@

format_cross_country <- function(crossmom, digits = 6, name="Euro") {
  rval <- format(round(crossmom, digits = digits), nsmall = digits)
  rval <- cbind(name,"", matrix(paste("$", rval, "$", sep = ""), ncol = 3))
  rval[1,1] <- name
  rval[2,1] <- ""
  rval[3,1] <- ""
  rval[1,2] <- "lower"
  rval[2,2] <- "medium"
  rval[3,2] <- "upper"
  rval <- apply(rval, 1, paste, collapse = " & ")
  rval <- paste(rval, "\\\\")
  rval[length(rval)] <- paste(rval[length(rval)], "\\hline")
  return(rval)
}

get_data <- function(i) {
  rval <- do.call("merge",
    lapply(i, function(j) results[[j]]$data))
  colnames(rval) <- names(results)[i]
  rval
}

@@

This \fixme{Jesus: sell method better, why this one and not another} framework allows us to assess the evolution of mean, variance and skewness over time. The skewness parameter turns out to be quite important to measure changes in the long run inflation dynamics, as it changes considerably over the time of the analysis. This trend would be all but invisible using a normal distribution.  

@@

The \fixme{Jesus: maybe comment on results here} remainder of this paper is structured as follows: Section \ref{s2} presents the data, Section \ref{s3} presents the model and the estimation techniques used, Section \ref{s4} illustrates our approach using Slovenia as an example, Section \ref{s5} presents the results and Section \ref{s6} concludes.

\clearpage
\section{Data}\label{s2}
%(... first method, than data...) \\
We \fixme{Jesus: yearly inflation rates better; don't start with data in paper} employ seasonally unadjusted HICP series of 21 countries ranging from January 1990 to December 2010 that are obtained from the \citet{oecd}. Countries included are Austria,	Belgium, Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,	Slovakia, Slovenia,  Spain,	Sweden, and	United Kingdom. Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity. Cypria and Malta are not included since they are very small economies. \\
The countries in this sample can be divided into three different groups. First the euro countries (Austria, Belgium, Estonia, France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, Netherlands, Portugal, Slovakia, Slovenia and Spain).  Second, the EU members not participating in the ERM~II: Czech Republic, Hungary, Poland, United Kingdom, and Sweden. Denmark stands on its own as a member of the EU and the ERM~II, but not yet a member of the EMU. \\
%(... schlussverkäufe schon hier erwähnen, einfach cool, ja das gibts, wir haben die aber weggeschätzt; generell viel angelsächsischer/präpotenter schreiben; nur 1 weg der richtige und zwar unserer; nicht erwähnen, dass andere dinge auch noch möglich; so nach motto: mich interessierts, ich machs, andere erst am ende erwähnen...)
% greece: -for winter discounts: First 10 days of January-end of February; -for summer discounts: First 10 days of July -end od August; but has always been included
As \readme{Jesus: Italy and Spain are problematic, fixed with same method as Belgium and Luxembourg} shown in a paper by \cite{pltm}, the harmonization of the treatment of sales prices led to considerable changes in some countries (Belgium, Italy, Luxembourg and Spain); failing to take this into account may lead to considerable distortions in the estimation. In these cases, we dealt with the inclusion of the sales periods effects\footnote{This was demanded by a Commission regulation, details may be found in: \citet{eurlex} and in the Appendix \ref{dataprobs}.} into the HICP series (January 2000 for Belgium, January 2001 for Italy, January 1999 for Luxembourg and January 2001 for Spain) by estimating dummies for these month\footnote{There are two month with a significant decrease in prices (January and July), followed by significant price increases in the following months (February and August).} and subtracting them from the original time series.

@@

To monitor changes in the dynamics of the inflation rates, we follow the usual strategy employed when testing for structural breaks. First, the model is fitted for the whole sample using the generalized logistic distribution. This then enables testing for the stability of the parameter estimates. If there is evidence of significant changes, the number and the timing of the break points are estimated. \\
This framework allows to trace the evolution of mean, variance and skewness of the inflation series and thus interpreting deviations in any of these parameters as possibly due to a change in monetary policy, is made feasible. Any huge intervention or change is captured by fluctuations in the parameter stability of the series and can be quantified and related to a distinct time frame. \\
%(...discuss correlation...) \\
The correlation \fixme{Jesus: yes, but a DGP needs to be assumed, it would need to take care of any autororrelation, wouldn't it?} dynamics of the series is of less importance to this framework, which concentrates on changes in mean, variance and skewness and does not focus on autocorrelation, the use of alternative tests, as described in \cite{andreou}, is also hindered by the low number of observations, due to the monthly frequencies.\footnote{In some years the Billion Prices Project of the \cite{bpp}, which provides a much higher data frequency, could solve this problem.} 
%To provide consistent estimates of the parameter vector $\hat{\phi}$, the covariance matrix $\hat{V}^{-1/2}$ is estimated using a heteroskedasticity and autocorrelation consistent covariance estimator.

@@

with location $\theta$, scale $\sigma$ and shape $\delta$. For $\delta=1$ the distribution simplifies to the logistic distribution, for $\delta<1$ it is skewed to the left and for $\delta$$>$1 it is skewed to the right (see Appendix \ref{glscore} for more details). \\
In econometrics, the logistic distribution is often used in income distributions and growth models. This is due to  its fatter tails, which fit these kind of data somewhat better. \citet{won} develop a maximum likelihood \fixme{Jesus: is this not better than what you do here: answer: no, they use completely other framework} estimator if the underlying distribution is a GL distribution. A similar GL distribution is also used in \citet{tolikas} who analyze extreme risk and value--at--risk in the German stock market, although they do not use a shape parameter. Regarding inflation rates, the GL distribution is --- to our best knowledge --- only used in context with expected inflation. \\
\citet{batchelor} use a logistic distribution (not its generalization) to model the distribution of mean expected inflation rates. \\

@@

\begin{eqnarray}
\mathit{efp}(t) ~=~ \hat{V}^{-1/2} n^{-1/2} \sum_{i=1}^{\lfloor nt \rfloor} s(y_i|\hat{\theta}, \hat{\sigma}, \hat{\delta}) \quad (0 \leq t \leq 1), \nonumber
\label{eqn2}
\end{eqnarray}
%(... jesus: thinks this is incomplete ...) \\
\fixme{Jesus: this seems a bit incomplete} where $\hat{V}$ is some consistent estimator of the variance of the scores, which is used to decorrelate the series as to avoid problems due to serial correlation. Then a functional central limit theorem (FCLT) for $\mathit{efp}(\cdot)$ is employed, which converges to a 3-dimensional Brownian bridge:

@@

<<echo=FALSE>>=
        x <- results$Slovenia
        slovpval <- sctest(x$efp, functional = supLM(0.1))$p.value
@

To illustrate our method, we consider the case of Slovenia. After a period of relative instability following the independence from Yugoslavia, Slovenia was successfull in realigning its economy and introduced a number of reforms leading to a stable growth in the past years. The good economic performance made it possible for Slovenia to enter the ERM~II in June 2004 and later on to introduce the euro in January 2007. The data collected contains the period from January 1996 onwards. In the first few years, Slovenia experienced very high inflation rates, but was successfull in reducing inflation to a much lower level. \\
A \fixme{Jesus: confidence interval for break} glance at the time series depicted in Figure \ref{fig1} of the monthly inflation rates reveals a decline in the mean around the middle of the observed time period. The $efp(\cdot)$ starts to cross the border as early as 2001 and reaches the peak somewhere between 2003 and 2004. The p--value of the supLM test is \Sexpr{round(slovpval, 5)} and clearly indicates at least one break.  The number of breaks is calculated by means of the LWZ criteria, which chooses one break in August 2003.\footnote{For the changes in the parameters, see Table \ref{resulttable}.} In this year, Slovenia introduced a number of financial reforms and severely decreased the year on year growth in the M3 monetary aggregate. Combined, these actions were enough to ensure the successfull participation in the ERM~II a year later and the introduction of the euro two years later. \readme{Jesus: include information on how the parameters change: I did give a footnote here}

@@

%(...3 punkte darstellung noch einbauen...) \\
%(...andere struktur: 1.short introduction, 2. countries in detail (6.1), 3. summary (6.2) ...) \\
%(... jesus: opening confusing and obscure...)\\
\readme{Jesus: include p-values in table; no place left; Jesus: opening seems a bit confusing and obscure} Table \ref{resulttable} presents the break dates and the parameter values of the segments as well as the dates of entry into the ERM/ERM~II and the date of the euro introduction. We formed seven groups of countries according to the observed pattern of break points. These seven groups consist of Austria and Belgium, the Eastern European countries, the two big southern European countries, Italy and Spain, Ireland, the countries with no change in the inflation dynamics, the countries with breaks in the early 90s and the member countries of the former DM--zone (excluding Austria and Belgium). \\

@@

\fixme{Jesus: this is just a data construct: but after desaisonalization still almost the same} Another combination of countries with rather identical results are the two biggest southern European countries, Italy and Spain. Both of them experienced high rates of inflation in the early 90s. Afterwards, in the phase leading up to the monetary union, mean and variance were greatly reduced. Soon after the fixing of the exchange rates, this trend changed significantly, leading to a higher mean and a much higher variance. A possible reason for this is that the fiscal responsibility that was required for an entry into the eurozone was softened after this goal was achieved. \\

@@

The disparity of the national inflation rates is still an issue. Many researchers, like \citet{hofmann}  find a considerable amount of inflation differentials. \citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity across EMU countries in terms of average inflation and its degree of persistence. In a paper examining structural convergence of inflation rates in EU countries, \citet{palomba} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar. They find that convergence over time of inflation dynamics was only partly observable. 
In a paper studying core inflation and using an aggregated euro area inflation rate, \citet{morana} finds three regimes (roughly 1980---1984, 1984---1993 and 1993---2000) governing the core inflation rate.  \\
Concerning inflation differentials, Figure \ref{greyplot} depicts the inflation rates and a rolling mean estimation, where the mean is calculated using the first moment of a fitted GL distribution.\footnote{If more countries have the same inflation rate in a given month, the lines are overlapped and thus get darker. We wold see one black line if all countries would show the same month--on--month increase at any one time. The red lines depict the means.} We do see that many of these countries follow a rather similar path. One event clearly visible is the financial crises and the short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months. Over time, there are hints, that the inflation rates converged towards some common -- albeit broad -- mean level. For the inflation series of the whole EMU, as published by the \citet{oecd}, we see \readme{Jesus: why not do this for EMU inflation (i.e. the one time series of the OECD} an increase in mean from \Sexpr{round(euhim1, 5)} (1990(1)--1998(12)) to \Sexpr{round(euhim2, 5)} (1999(1)--2008(8)) with a strong decrease to \Sexpr{round(euhim3, 5)} (2008(9)--2010(12)) following the financial crises. Whereas variance has increased over time from    \Sexpr{round(euhiv1, 5)} to \Sexpr{round(euhim3, 5)}. Regarding skewness, we see a clear change towards a negative skewness from \Sexpr{round(euhis1, 5)} to \Sexpr{round(euhis3, 5)}.

@@

Regarding the general trend, Table \ref{grouptable} presents lower, medium and upper (25, 50, 75) percentiles of the first three moments of the inflation series  at the beginning of the convergence phase (January 1995), the beginning of the financial crisis (in September 2008\footnote{After the insolvency of the Lehman Brothers in September 2008.}) and for the last observation in the sample. The countries are grouped into the ``core'' of the eurozone (Austria, Belgium, Finland, France, Germany, Luxembourg and Netherlands), the PIIGS countries (Portugal, Italy, Ireland, Greece and Spain) and all the others.  \\

\begin{longtable}{|l|l|l|l|l|}
\hline
\multicolumn{5}{|c|}{1995} \\
\hline
Zone & Percentile & Mean & Variance & Skewness\\
\hline
<<echo=FALSE, results=tex>>=
euro <- c(1, 2, 6, 7, 8, 13, 14)
noneuro <- c(3, 4, 5, 10, 15, 17, 18, 20, 21)
piigs <- c(9, 11, 12, 16, 19)
writeLines(format_cross_country(cross_country_moments(euro, 1995), name ="Euro"))
@
\hline
<<echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(piigs, 1995), name ="PIIGS"))
@
\hline
<<echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(noneuro, 1995), name ="Noneuro"))
@
\hline
\multicolumn{5}{|c|}{2008(9)} \\
\hline
<<echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(euro, 2008+8/12), name ="Euro"))
@
\hline
<<echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(piigs, 2008+8/12), name ="PIIGS"))
@
\hline
<<echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(noneuro, 2008+8/12), name ="Noneuro"))
@
\hline
\multicolumn{5}{|c|}{2010(12)} \\
\hline
<<echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(euro, 2010+11/12), name ="Euro"))
@
\hline
<<echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(piigs, 2010+11/12), name ="PIIGS"))
@
\hline
<<echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(noneuro, 2010+11/12), name ="Noneuro"))
@
\hline
\caption{Development of first three moments}
\label{grouptable}
\end{longtable}

Regarding the mean, there is a slight increase in median inflation in the euro countries, with a bigger increase in the PIIGS countries, whereas the other countries witness a strong decrease in inflation (as a result of their efforts to reach the qualifications required for eurozone entry). The variance for the euro countries almost triples over the past 15 years, whereas the variance of the PIIGS countries decreases slightly. Regarding skewness, in the euro countries and the PIIGS, there is a tendency of decreasing skewness, indicating a shift of the distribution to the right, whereas the  countries outside the euro do not show much change.\readme{this way, there is a completely different story, not the big changes that we get if we take whole subperiods in the data.}
